---

title: "robin"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ robin}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
author: "Valeria Policastro - Dario Righelli"

---

In network analysis, many community detection algorithms have been developed. 
However,their applications leave unaddressed one important question: 
the statistical validation of the results. 
Are the detected communities significant or are they a result of chance only 
due to the positions of edges in the network?

ROBIN (ROBustness In Network) is an R package for the validation of community 
detection it has a double aim it studies the robustness of a community 
detection algorithm and compares the robustness of two community algorithms. 

The package implements a methodology that detects if the community structure 
found by a detection algorithm is statistically significant or is a result 
of chance, merely due to edge positions in the network. It performs a 
perturbation strategy and runs a null model to build a set of procedures based 
on different stability measures. 

In particular, it provides:
1)A procedure to examine the stability of the partition recovered against random 
perturbations of the original graph structure
2)Three tests to determine whether the obtained clustering departs significantly 
from the null model
3)A routine to compare different detection algorithms applied to the same 
network to discover which fits better
4)A graphical interactive representation 

# Install package
```{r message=FALSE, warning=FALSE, paged.print=TRUE}
library(devtools)
install_github("ValeriaPolicastro/robin",force=TRUE)
library("robin")
```

# Preparation of the Graph 
As input, the **ROBIN** package expects a network that can be read from different 
format: edgelist, pajek, graphml, gml, ncol, lgl, dimacs, graphdb and igraph 
graphs.

With the **prepGraph** function we create an *igraph* object, from the selected 
input file, needed for the ROBIN execution.

```{r}
my_network <- system.file("example/football.gml", package="robin")
# downloaded from: http://www-personal.umich.edu/~mejn/netdata/
graph <- prepGraph(file=my_network, file.format="gml")
graph
```

# Random Graph 
In such a way, we can create the Random graph (null model) that has to be used 
for the package, the graph argument must be of the same type returned by the 
**prepGraph** function.

```{r}
graphRandom <- random(graph=graph)
graphRandom
```

# Plot graph
To have a graphical representation of the network we use the **plotGraph** 
function implemented with aid of *networkD3* package.

```{r}
plotGraph(graph)
```

# Create Community 
To create the communities, in the package are implemented all the igraph 
algorithms.

```{r}
methodCommunity(graph=graph, method="fastGreedy") #as community 
membershipCommunities(graph=graph, method="fastGreedy") #as membership
```

# Community Plot

It gives an interactive 3D plot of the communites for the chosen algorithm.

```{r}
members <- membershipCommunities(graph=graph, method="fastGreedy")
plotComm(graph=graph, members=members)
```

# Procedure to validate the robustness
It creates the ROBIN procedure to validate the robustness of your network.
In this example we used the "vi" distance as a stability measure, the indipendent 
type of procedure and the louvain algorithm as a community detection algorithm, 
but users can choose different measures ("nmi","split.join", "adjusted.rand") and 
algorithms (walktrap", "edgeBetweenness", "fastGreedy", "spinglass", "leadingEigen",
"labelProp", "infomap", "optimal", "other") implemented in the package.
To make things easier we save the output list in the *proc* variable to use it 
later.

```{r}
proc <- robinRobust(graph=graph, graphRandom=graphRandom, measure="vi", 
                  method="louvain", type="independent")
```

# Robin Plots
Now in order to compare the obtained curves from the Procedure we can make a 
plot with the **plotRobin** function.
It shows on the y-axis the average of the chosen measure while on the x-axis 
the percentuage of perturbation of the real data and the null model.

The *model1* and *model2* arguments are, respectively, the *Mean* and the 
*MeanRandom* contained inside the **robinRobust** output list.

```{r}
plotRobin(graph=graph, model1=proc$Mean, model2=proc$MeanRandom, 
legend=c("real data", "null model"), measure="vi")
```

# Statistical Tests between Real data and Null model 
Now we test the differeces between this two curves with:
- Functional data analysis
- Gaussian Process
- Area Under the Curve (AUC)
The *model1* and *model2* arguments are, respectively, the *Mean* and the 
*MeanRandom* contained inside the **robinRobust** output list.

```{r}
robinFDATest(graph=graph, model1=proc$Mean, model2=proc$MeanRandom, 
             measure="vi")
robinGPTest(ratio=proc$ratios)
robinAUC(graph=graph, model1=proc$Mean, model2=proc$MeanRandom, 
             measure="vi")
```

# Comparison Two different Methods
Now we want to see which algorithm (between two of them) better detects the 
communities.
For example if we take the *Fast Greedy* and *Louvain* algorithms, 
we want to check which one better fits our network.
We firstly plot them with their different communities to have an idea of 
the network of interest.

```{r}
membersFast <- membershipCommunities(graph=graph, method="fastGreedy")
membersLouv <- membershipCommunities(graph=graph, method="louvain")
plotComm(graph=graph, members=membersFast)
plotComm(graph=graph, members=membersLouv)
```

Secondly, we run the **robinCompare** function to compare the two algorithms, 
which, as before, we store the output in the *comp* variable to use it later.

```{r}
comp <- robinCompare(graph=graph, method1="fastGreedy",
                method2="louvain", measure="vi", type="independent")
```

Thirdly, we plot the two curves of the two methods in comparison.
The *model1*, *model2* arguments are, respectively, 
the *Mean1*, *Mean2* contained inside the **robinCompare** output list.

```{r}
plotRobin(graph=graph, model1=comp$Mean1, model2=comp$Mean2, measure="vi", 
legend=c("fastGreedy", "louvain"), title="FastGreedy vs Louvain")
```
In this example, the Louvain algorithm fits better the network of interest, as 
the found communities varies less than the Fast greedy method.

# Statistical Tests between two community detection algorithms
Now we test the differeces between the two curves, as we did before but for the
comparison of the two different methods the *model1* argument must be the 
*Mean1* and the *model2* argument must be the *Mean2* both contained inside 
the **robinCompare** output list.

```{r}
robinFDATest(graph=graph, model1=comp$Mean1, model2=comp$Mean2, measure="vi")
robinGPTest(ratio=comp$ratios1vs2)
robinAUC(graph=graph, model1=comp$Mean1, model2=comp$Mean2, measure="vi")
```

