---
title: "robin"
subtitle: "ROBustness In Network"
author: "Valeria Policastro"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{robin}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  knitr:::html_vignette:
    toc: yes
    toc_depth: 2
---


# robin

In network analysis, many community detection algorithms have been developed. 
However,their applications leave unaddressed one important question: 
**the statistical validation of the results**. 


*robin* (ROBustness in Network) is an R package for the validation of community 
detection. It has a double aim: **to study the robustness** of a community 
detection algorithm and **to compare** the robustness of **different community detection algorithms** on the same network. 

It provides:
1) a procedure to examine the robustness of a community detection algorithm against random 
perturbations of the original graph; 
2) two tests to determine the statistical difference between stability measure curves;
3) a routine to choose among different community detection algorithms the one that better fits the network of interest;
4) a graphical interactive representation.

# Loading package
```{r message=FALSE, warning=FALSE, paged.print=TRUE}
library("robin")
```

# Input 
The input of the *robin* package is a network that can be read from different 
format: edgelist, pajek, graphml, gml, ncol, lgl, dimacs, graphdb and igraph 
graphs.

**prepGraph** function creates an *igraph* object from the input file. This step is necessary for *robin* execution

```{r}
my_network <- system.file("example/football.gml", package="robin")
# downloaded from: http://www-personal.umich.edu/~mejn/netdata/
graph <- prepGraph(file=my_network, file.format="gml")
graph
```

# Null model 

*robin* offers two choiches for the null model:

1) external building according to users’ preferences, then the null graph is passed as a
variable;

2) generation by using the function **random**.

**random** function creates the null model for the robustness procedure **robinRobust**. The graph argument must be the same returned by **prepGraph** function. 

The function **random** creates a random graph with the same degree distribution of the
original graph, but with completely random edges, using the **rewire** function of the package
igraph with the option "keeping_degseq" that preserves the degree distribution of the original
network. The function **rewire** randomly assigns a number of edges between vertices with
the given degree distribution.

```{r}
graphRandom <- random(graph=graph)
graphRandom
```

# Network visualization
**plotGraph** function offers a graphical representation of the network with the aid of *networkD3* package.

```{r}
plotGraph(graph)
```

## Community detection 
**methodCommunity** function detects communities using all the algorithms implemented in *igraph* package:  "walktrap", "edgeBetweenness", "fastGreedy", "spinglass", "leadingEigen",
"labelProp", "infomap", "optimal", "other".

```{r}
methodCommunity(graph=graph, method="fastGreedy") 
```
**membershipCommunities** function detects the community membership.
```{r}
membershipCommunities(graph=graph, method="fastGreedy") 
```

## Community visualization

**plotComm** function produces an interactive 3D plot of the communites detected by the chosen algorithm.

```{r}
members <- membershipCommunities(graph=graph, method="fastGreedy")
plotComm(graph=graph, members=members)
```

# Robustness of a community detection algorithm
**robinRobust** function implements the validation of the community robustness.
In this example we use "vi" distance as stability measure, "indipendent" 
type procedure and "louvain" as community detection algorithm. 

Users can choose different measures: "nmi","split.join", "adjusted.rand".

```{r}
proc <- robinRobust(graph=graph, graphRandom=graphRandom, measure="vi", 
                  method="louvain", type="independent")
```

**plotRobin** function plots the stability measure curves. 
The (x,y)-axes represent the percentuage of perturbation and the average of the stability measure, respectively.

```{r}
plotRobin(graph=graph, model1=proc$Mean, model2=proc$MeanRandom, 
legend=c("real data", "null model"), measure="vi")
```

## Statistical tests
The differeces between the stability measure curves are tested using:
1) Functional Data Analysis (FDA);
2) Gaussian Process (GP).
Moreover the Area Under the Curve (AUC) is evaluated too.

**robinFDATest** function implements the FDA testing.

```{r}
robinFDATest(graph=graph, model1=proc$Mean, model2=proc$MeanRandom, 
             measure="vi")
```

The first figure represents the stability measure plot using "louvain" algorithm for detecting communities. The second one represents the corresponding adjusted p-values of the Interval Testing procedure. Horizontal red line corresponds to the critical value 0.05. 

**robinGPTest** function implements the GP testing.
```{r message=FALSE, warning=FALSE, paged.print=TRUE}
robinGPTest(model1=proc$Mean, model2=proc$MeanRandom)
```
The output is the Bayes Factor.

**robinAUC** function implements the AUC.

```{r}
robinAUC(graph=graph, model1=proc$Mean, model2=proc$MeanRandom, 
             measure="vi")
```
The outputs are the area under the two curves


# Comparison of two community detection algorithms

**robinCompare** function compares two detection algorithms on the same network  and permitts the user
to choose the one that better fits the network of interest.

In this example we consider the "Fast Greedy" and "Louvain" algorithms.

We firstly plot the communities dectected by both algorithms.

```{r}
membersFast <- membershipCommunities(graph=graph, method="fastGreedy")
membersLouv <- membershipCommunities(graph=graph, method="louvain")
plotComm(graph=graph, members=membersFast)
plotComm(graph=graph, members=membersLouv)
```

Secondly, we compare them with **robinCompare** function. 


```{r}
comp <- robinCompare(graph=graph, method1="fastGreedy",
                method2="louvain", measure="vi", type="independent")
```

Thirdly, we plot the curves of the compared methods.


```{r}
plotRobin(graph=graph, model1=comp$Mean1, model2=comp$Mean2, measure="vi", 
legend=c("fastGreedy", "louvain"), title="FastGreedy vs Louvain")
```

In this example, the Louvain algorithm fits better the network of interest, 
as the curve of the stability measure assumes lower values than the one obtained by 
the Fast greedy method. 

## Statistical tests

The following procedures test the statistical differeces between the two curves using two different methods
```{r}
robinFDATest(graph=graph, model1=comp$Mean1, model2=comp$Mean2, measure="vi")

```

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
robinGPTest(model1=comp$Mean1, model2=comp$Mean2)
```

while 
```{r}
robinAUC(graph=graph, model1=comp$Mean1, model2=comp$Mean2, measure="vi")
```
calculates the area under the curves.
