---
title: "Vignette"
author: "Valeria Policastro - Dario Righelli"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: true
  pdf_document: 
    toc: true
editor_options: 
  chunk_output_type: console
---

#Abstract
In network analysis, many community detection algorithms have been developed. 
However,their applications leave unaddressed one important question: 
the statistical validation of the results. 
Are the detected communities significant or are they a result of chance only 
due to the positions of edges in the network?

ROBIN (Robustness In Network) gives a statistical answer to the validation of 
the community structure by looking at the robustness of the network.

The package implements a methodology that detects if the community structure 
found by a detection algorithm is statistically significant or is a result 
of chance, merely due to edge positions in the network. It performs a 
perturbation strategy and runs a null model to build a set of procedures based 
on the Variation of Information as a clustering distance. 

In particular, it provides:
1)A procedure to examine the stability of the partition recovered against random 
perturbations of the original graph structure
2)Three tests to determine whether the obtained clustering departs significantly 
from the null model
3)A routine to compare different detection algorithms applied to the same 
network to discover which fits better
4)A graphical interactive representation 

### Install packages
```{r}
# install.packages("devtools")
library(devtools)
install_github("ValeriaPolicastro/robin")
library("robin")
library("cowplot")
library("gprege")
```


# Preparation of the Graph 
As input, the **ROBIN** package expects a network that can be read from different 
format: edgelist, pajek, graphml, gml, ncol, lgl, dimacs, graphdb and igraph 
graphs.

With the **prepGraph** function we create an *igraph* object, from the selected 
input file, needed for the ROBIN execution.

```{r}
my_file <- system.file("example/football.gml", package="robin")
# downloaded from: http://www-personal.umich.edu/~mejn/netdata/
graph <- prepGraph(file=my_file, file.format="gml")
graph
```

# Random Graph 
In such a way, we can create the Random graph (null model) that has to be used 
for the package, the graph argument must be of the same type returned by the 
**prepGraph** function.

```{r}
graphRandom <- random(graph=graph)
```

# Plot graph
To have a graphical representation of the network we use the **plotGraph** 
function implemented with aid of *networkD3* package.

```{r}
plotGraph(graph)
```

# Create Community 
To create the communities, in the package are implemented all the igraph 
algorithms.

```{r}
methodCommunity(graph=graph, method="fastGreedy") #as community 
membershipCommunities(graph=graph, method="fastGreedy") #as membership
```

#Community Plot

It gives an interactive 3D plot of the communites for the chosen algorithm.

```{r}
plotCommu(graph=graph, method="fastGreedy")
```

# Procedure to validate the robustness
It creates the ROBIN procedure to validate the robustness of your network.
To make things easier we save the output list in the *proc* variable because 
to use it later.

```{r}
proc <- robinProc(graph=graph, graphRandom=graphRandom, method="louvain",
                 type="independent")
```

# Robin Plots
Now in order to compare the obtained curves from the Procedure we can make a 
plot with the **plotRobin** function.
It shows on the y-axis the average of the Variation of Information (VI) measure 
while on the x-axis the percentuage of perturbation of the real data and 
the null model.

The *model* and *modelR* arguments are, respectively, the *viMean* and the 
*viMeanRandom* contained inside the *robinProc* output list.

```{r}
plotRobin(graph=graph, model=proc$viMean, modelR=proc$viMeanRandom)
```

# Statistical Tests between Real data and Null model 
Now we test the differeces between this two curves with:
- Functional data analysis
- Gaussian Process
- Area Under the Curve (AUC)
The *model1* and *model2* arguments are, respectively, the *viMean* and the 
*viMeanRandom* contained inside the *robinProc* output list.

```{r}
robinFDATest(graph=graph, model1=proc$viMean, model2=proc$viMeanRandom)
robinGPTest(ratio=proc$ratios)
robinAUCTest(graph=graph, model1=proc$viMean, model2=proc$viMeanRandom)
```

# Comparison Two different Methods
Now we want to see which algorithm (between two of them) better detects the 
communities.
For example if we take the *Fast Greedy* and *Louvain* algorithms, 
we want to check which one better fits our network.
We firstly plot them with their different communities to have an idea of 
the network of interest.

```{r}
plotCommu(graph=graph, method="fastGreedy")
plotCommu(graph=graph, method="louvain")
```

Secondly, we run the **comparison** function to compare the two algorithms, 
which, as before, we store the output in the *comp* variable to use it later.

```{r}
comp <- comparison(graph=graph, graphRandom=graphRandom, method1="fastGreedy",
                method2="louvain",type="independent")
```

Thirdly, we plot the curves of the null model and the real data for both
methods and the two curves of the two methods in comparison.
The model1 argument must be the viMean1 output of the comparison function, the 
modelR1 argument must be the viMeanRandom1,the model2 argument must be the 
viMean2 and the modelR2 argument must be the viMeanRandom2 all outputs of the
comparison function.

```{r}
plotRobinCompare(graph=graph, model1=comp$viMean1, model2=comp$viMean2,
                 modelR1=comp$viMeanRandom1, modelR2=comp$viMeanRandom2,
                 legend=c("real data", "null model"),
                 legend1vs2=c("Fast Greedy","Louvain"),
                 title1="Fast Greedy", title2="Louvain",
                 title1vs2="Fast Greedy vs Louvain")
```

# Statistical Tests between two methods 

We then test them, as we did it, before but for the comparison the model1 
argument must be the viMean1 output of the function comparison and the model2 
argument must be the viMean2 output of the function comparison.

```{r}
robinFDATest(graph=graph, model1=comp$viMean1, model2=comp$viMean2)
robinGPTest(ratio=comp$ratios1vs2)
robinAUCTest(graph=graph, model1=comp$viMean1, model2=comp$viMean2)
```

